---
title: "social_atlas_code"
author: "Matthew Lam"
date: '2022-03-31'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
#load libraries
library(tidyverse)
library(tmap)
library(sf)
library(sp)
library(rgdal)
library(rgeos)
library(janitor)
library(RColorBrewer)
library(spdep)

#working directory
setwd("~/Library/CloudStorage/OneDrive-UniversityCollegeLondon/Y2/Term 2/Geocomputation/social-atlas/social-atlas")

#read in the data for vaccines
wc15apr <- read.csv("./vax raw data/wc15apr.csv")

#read in the data for case rates
london_case_rates <- read.csv("msoa_case_rates.csv")

#read in the data for ethnicity
ethnicitiesdata <- read.csv("msoa_ethnicities_data.csv")

#read in the data for population numbers 2011
population_numbers <- read.csv("2011_population_numbers.csv")

#read in the data for population numbers 2020
population_numbers_2020 <- read.csv("2020_population_numbers.csv")

#read in the data for qualification numbers
qualification_numbers <- read.csv("qualification_numbers.csv")

#read in the data for imd decile by lsoa
lsoa_imd_decile <- read.csv("deprivation_lsoa.csv")

#read in the lsoa to msoa lookup file (used later to summarise imd by msoa)
msoa_lsoa_lookup <- read.csv("msoa_lsoa_lookup.csv")

#read in the data for vax sites
pharmacies <- read.csv("pharmacies_vax_sites.csv")
gps <- read.csv("gp_led_vax_sites.csv")
hospitals <- read.csv("hospital_vax_sites.csv")
```

## Data Cleaning

#### Week Commencing 13 May 2020 Vaccinations
```{r}
#subset for london only
wc13may_london <- wc13may %>% filter(region_name == "London")

#totals for first dose
wc13may_london$total_firstdose <- wc13may_london$under40_firstdose + wc13may_london$X40to44_firstdose + wc13may_london$X45to49_firstdose + wc13may_london$X50to54_firstdose + wc13may_london$X55to59_firstdose + wc13may_london$X60to64_firstdose + wc13may_london$X65to69_firstdose + wc13may_london$X70to74_firstdose + wc13may_london$X75to79_firstdose + wc13may_london$X80plus_firstdose

#select only the relevant columns 
london_vax_rates13may <- wc13may_london %>% dplyr::select(region_name, ltla_name, msoa_code, total_firstdose)

#expressing as a proportion of total msoa population
#first merge the data
london_vax_rates13may <- merge(london_vax_rates13may, population_numbers_2020, by.x="msoa_code", by.y="msoa_code")

#calculating the percentage population vaccinated 
london_vax_rates13may$per_atleastfirstdose <-(london_vax_rates13may$total_firstdose / london_vax_rates13may$pop) *100

#round
london_vax_rates13may$per_atleastfirstdose <- floor(london_vax_rates13may$per_atleastfirstdose)
```

#### Qualifications

```{r}
#merging the dataset with population numbers
qualification_msoa <- merge(population_numbers, qualification_numbers, by.x="msoa_code", by.y="msoa_code")

#calculating percent no qualification 
qualification_msoa$percent_noqual <- (qualification_msoa$persons_noqualification / qualification_msoa$pop)*100

#calculating percent degreeandabove
qualification_msoa$percent_degreeandabove <- (qualification_msoa$persons_degree_and_above / qualification_msoa$pop)*100

```

#### Deprivation

```{r}
#first we need to merge the lsoaimd file with the lookup file 
lsoa_imd_lookup <- merge(lsoa_imd_decile, msoa_lsoa_lookup, by.x="lsoa_code", by.y="LSOA11CD")

#inspect
View(lsoa_imd_lookup)

#next we need to summarise by msoa. First we will group it by code and name, and then summarise by the mean imd of all the LSOAs in that MSOA. 
imd_decile_msoa <- lsoa_imd_lookup %>% group_by(MSOA11CD, MSOA11NM) %>% summarise(mean_imd_decile= mean(imd_decile))
View(imd_decile_msoa) 


#remove city of london by subsetting for everything that is not city of london
imd_decile_msoa <- subset(imd_decile_msoa, MSOA11CD != "E02000001")

#Round up the mean imd decile
imd_decile_msoa$mean_imd_decile <- round(imd_decile_msoa$mean_imd_decile, digits=0)

#change to a numeric variable
imd_decile_msoa$mean_imd_decile <- as.numeric(imd_decile_msoa$mean_imd_decile)

#inspect
View(imd_decile_msoa)

```

#### Vax Sites
```{r}
#filter for London Only
hospitals <- hospitals %>% filter(Region == "London")
pharmacies <- pharmacies %>% filter(Region=="London")
gps <-gps %>% filter(Region == "London")

#select only postcode column
hospitals <- hospitals[,5:6]
pharmacies <- pharmacies[,6:7]
gps <- gps[,6:7]


#bind all together to one column
allsites_raw <- rbind(hospitals, pharmacies, gps)
View(allsites_raw)

#writw as new csv for use in postcode to coordinate finder
write.csv(allsites_raw, "allsites_raw.csv")
```




## Vaccines Map

#### Shapefile

```{r}
#msoa load the shape
london_msoa_shape <- read_sf("./ESRI/MSOA_2011_London_gen_MHW.shp")
plot(london_msoa_shape)
```

#### Vaccines Map

```{r}

# change data types
london_vax_rates13may  <- mutate_at(vaccines_msoa_sdf13may , vars(per_atleastfirstdose),
    function(x) as.numeric(x))

#histogram to inspect data
hist(london_vax_rates13may$per_atleastfirstdose)

#joining to the spatial dataframe
vaccines_msoa_sdf13may <- left_join(london_msoa_shape,london_vax_rates13may, by = c("MSOA11CD" = "msoa_code"))

#defining a custom palette using cartocolors colors
mypal1 <- c("#f9ddda","#f2b9c4","#e597b9","#ce78b3","#ad5fad","#834ba0","#573b88")

#vaccines map 
vaccines_map <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(vaccines_msoa_sdf13may)+
  tm_fill(col="per_atleastfirstdose", n=5,
              style="jenks", title="% Population with At Least \nOne Dose", 
              palette=mypal1, 
              border.col="white")+
  tm_borders(col="grey", lwd=0.3)+
  tm_layout(main.title="Population in London with At Least One Dose \n (As of 13 May 2021)",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame = F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))
vaccines_map

#save the map as a pdf
tmap_save(vaccines_map, "vaccines_map.pdf")

```

## Hotspot Analysis for Vaccines

```{r}
#define neighbours
msoa_neighbours_fd <- dnearneigh(st_geometry(st_centroid(vaccines_msoa_sdf13may)), 0, 3000)


# creates a neighbours list
msoa_spatial_weights_fd <- msoa_neighbours_fd %>%
    nb2listw(., style = "B")


# run the local Gi* test on our  data, note we again use the fd
# weights
vaccines_LGO <- vaccines_msoa_sdf13may %>%
    pull(per_atleastfirstdose) %>%
    as.vector() %>%
    localG(., msoa_spatial_weights_fd)

# print the local Gi* statistic
head(vaccines_LGO)

# join the local Gi* statistic to `obesity_ward_sdf` spatial dataframe
vaccines_msoa_sdf <- vaccines_msoa_sdf %>%
    mutate(vaccines_LGO_G = as.numeric(vaccines_LGO))

# create a colour palette
GIColours <- rev(brewer.pal(8, "RdBu"))

#subset london boroughs to include outline
ldnb <- read_sf("./ESRI/London_Borough_Excluding_MHW.shp")
towerhamlets_outline <- subset(ldnb, NAME=="Tower Hamlets")
westminster_outline <- subset(ldnb, NAME=="Westminster")
hackney_outline <- subset(ldnb, NAME=="Hackney")
camden_outline <- subset(ldnb, NAME=="Camden")

# plot the clusters
vax_hotspot <- tm_shape(vaccines_msoa_sdf) + 
  tm_polygons("vaccines_LGO_G", style = "pretty", palette=GIColours,
              midpoint=0, title = "Local Gi* statistic") +
  tm_layout(main.title = 'Hot/Cold Spot Map of COVID-19 Vaccination Rates in London', 
            main.title.fontface = 2, fontfamily = "Helvetica",
            legend.outside = TRUE, 
            legend.outside.position = "right", 
            legend.title.size = 1, 
            legend.title.fontface = 2, 
            frame=F) + 
  tm_compass(type = "arrow", position = c("right", "bottom")) +
  tm_scale_bar(breaks = c(0, 5, 10, 15, 20), position = c("left", "bottom"))+
  tm_shape(towerhamlets_outline)+
  tm_borders(col="yellow")+
  tm_shape(westminster_outline)+
  tm_borders(col="yellow")+
  tm_shape(hackney_outline)+
  tm_borders(col="yellow")

#save as pdf
tmap_save(vax_hotspot, "vax_hotspot.pdf")
```

## Global Moran's I

```{r}
#joining to the spatial dataframe
vaccines_msoa_sdf13may <- left_join(london_msoa_shape,london_vax_rates13may, by = c("MSOA11CD" = "msoa_code"))

# remove NAs
vaccines_msoa_sdf13may <- vaccines_msoa_sdf13may %>%
    filter(!is.na(per_atleastfirstdose))

# calculate the centroids of all of the msoas in London
msoa_centroid <- vaccines_msoa_sdf13may %>%
    st_centroid()

# create a neighbours list based on the Queen neighbour definition
msoa_neighbours_queen <- vaccines_msoa_sdf13may %>%
    poly2nb(., queen = T, snap=10)

# creates a neighbours list: Fixed Distance
msoa_spatial_weights_fd <- msoa_neighbours_fd %>%
    nb2listw(., style = "B")

# plot neighbours: Queen
plot(msoa_neighbours_fd, st_geometry(msoa_centroid), col = "red", pch = 20, cex = 0.5)


#Queen's neighbour list
msoa_spatial_weights_queen <- msoa_neighbours_queen %>% 
  nb2listw(., style="C")

#plot
vaccines_MI_sp <- moran.plot(vaccines_msoa_sdf13may$per_atleastfirstdose, listw = msoa_spatial_weights_queen)


# run the global Moran's I test 
vaccines_GMI <- vaccines_msoa_sdf13may %>%
    pull(per_atleastfirstdose) %>%
    as.vector() %>%
    moran.test(., msoa_spatial_weights_queen)

# print global Moran's I result
vaccines_GMI

#export table using stargazer
library(stargazer)
stargazer(vaccines_GMI)
```



## Local Moran's I

```{r}


# run the local Moran's I test 
vax_LMI <- vaccines_msoa_sdf13may  %>%
    pull(per_atleastfirstdose) %>%
    as.vector() %>%
    localmoran(.,msoa_spatial_weights_queen)

# coerce into dataframe
vax_LMI <- as.data.frame(vax_LMI)

# update the names for easier reference
names(vax_LMI) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")

# join to obesity_ward_sdf
vaccines_msoa_sdf13may <- vaccines_msoa_sdf13may %>%
    mutate(vax_LMI)

# significance breaks
breaks <- c(0, 0.05, 0.1, 1)

# colours for our local Moran's I statistic
colours <- c('white', "#a6bddb", "#2b8cbe" )

# map our local Moran's I statistic 
tm_shape(vaccines_msoa_sdf13may) + 
  tm_polygons("LMI_sigP", style = "fixed", breaks = breaks, 
              palette = rev(colours), title = "p-value of Local Moran's I stat") + 
  tm_borders(col="grey", lwd=0.3)+
  tm_layout(main.title = "Spatial Autocorrelation of \nVaccination Rates in London", 
            main.title.fontface = 2, fontfamily = "Helvetica", 
            legend.outside = TRUE, 
            legend.outside.position = "right", 
            legend.title.size = 1, 
            legend.title.fontface = 2, frame=F) + 
  tm_compass(type = "arrow", position = c("right", "bottom")) + 
  tm_scale_bar(breaks = c(0, 5, 10, 15, 20), position = c("left", "bottom"))


# compare local value for obesity with mean value
vaccines_msoa_sdf13may <- vaccines_msoa_sdf13may %>% mutate(obs_diff = (per_atleastfirstdose - mean(vaccines_msoa_sdf13may$per_atleastfirstdose)))

# compare local LMI value with mean LMI value
vaccines_msoa_sdf13may <- vaccines_msoa_sdf13may %>% mutate(LMI_diff = (vax_LMI$LMI_Ii - mean(vax_LMI$LMI_Ii)))

# set a significance threshold 
signif <- 0.05

# generate column with cluster type, using values above
vaccines_msoa_sdf13may <- vaccines_msoa_sdf13may %>% 
  mutate(cluster_type = case_when(obs_diff > 0 & LMI_diff > 0 & LMI_sigP < signif ~ "High-High",
                                  obs_diff < 0 & LMI_diff < 0 & LMI_sigP < signif ~ "Low-Low",
                                  obs_diff < 0 & LMI_diff > 0 & LMI_sigP < signif ~ "Low-High",
                                  obs_diff > 0 & LMI_diff < 0 & LMI_sigP < signif ~ "High-Low",
                                  LMI_sigP > signif ~ "No Significance"))


# count the different values within our `cluster_type` variable
count(vaccines_msoa_sdf13may, cluster_type)

# create a colour palette
pal <- c("#d7191c", "#fdae61", "#abd9e9", "#2c7bb6","#F5F5F5")

# plot the clusters
lisamap <- tm_shape(vaccines_msoa_sdf13may) +
  tm_fill(col = "cluster_type", palette = pal, title = "Cluster Type") +
  tm_borders(col="grey", lwd=.7)+
  tm_layout(main.title = "Cluster Map of Vaccination Rates in London",
            main.title.fontface = 2, fontfamily = "Helvetica", legend.outside = TRUE,
            legend.outside.position = "right",
            legend.title.size = 1,
            legend.title.fontface = 2, 
            frame=F) +
    tm_shape(towerhamlets_outline)+
  tm_borders(col="black")+
  tm_shape(westminster_outline)+
  tm_borders(col="black")+
  tm_shape(hackney_outline)+
  tm_borders(col="black")+
  tm_compass(type = "arrow", position = c("right", "bottom")) +
  tm_scale_bar(breaks = c(0, 5, 10, 15, 20), position = c("left", "bottom"))

#save as pdf
tmap_save(lisamap, "lisamap.pdf")
```

## Breakdown by Borough Graph

```{r}
#vax totals in each borough
vax_by_borough <- london_vax_rates13may %>% group_by(lad_name) %>% summarise(total_firstdose = sum(total_firstdose), pop = sum(pop))
View(vax_by_borough)

#vax totals in each borough divide by population 
vax_by_borough$per_atleastfirstdose <- (vax_by_borough$total_firstdose / vax_by_borough$pop)*100

#round to nearestinteger
vax_by_borough$per_atleastfirstdose <- floor(vax_by_borough$per_atleastfirstdose)


#make a bar graph ordered by vax rate per borough
plot1 <- vax_by_borough %>% 
  mutate(lad_name = fct_reorder(lad_name, desc(per_atleastfirstdose))) %>%
  ggplot(aes(x=lad_name, y=per_atleastfirstdose))+
  geom_bar(stat="identity", fill="#f68060")+
  coord_flip()+
  xlab("")+
  ylab("Percent of Population Vaccinated with At Least One Dose")+
  ggtitle("Breakdown of Vaccination Rates by London Boroughs")

#save as pdf
ggsave("plot1.pdf",plot1)
    
```

## Black and Black British Map

```{r}
#using tmap
black <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
  tm_shape(ethnicity_msoa_sdf)+
  tm_fill(c("per_black"), 
            style="jenks", 
            palette=mypal2, 
            title=c("% Population Black/Black British"))+
  tm_borders(col="grey", lwd=0.3)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))+
  tm_layout(main.title="Distribution of Black/Black British \nPopulations in London",
            main.title.fontface=2,
            fontfamily="Helvetica", legend.outside=T, legend.outside.position = c("right", "bottom"), frame=F)

#save as pdf
tmap_save(black, "black.pdf")
```



## Young People

```{r}
#read in the data
youth <- read.csv("youthdata.csv")

#calculate proportion of population aged 16 to 29 based on 2011 estimates
youth$per_16to29 <- (youth$nr_16to19 / youth$X2011_midyear_estimate)*100

write.csv(youth, "youthdata_msoa.csv")

#map
#joining to the spatial dataframe
youth_msoa_sdf <- left_join(london_msoa_shape, youth, by = c("MSOA11CD" = "msoa_code"))

#defining a custom palette using cartocolors colors
mypal9 <- c("#ecda9a","#efc47e","#f3ad6a","#f7945d","#f97b57","#f66356","#ee4d5a")

#youth map
youth <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(youth_msoa_sdf)+
  tm_fill(col="per_16to29", n=5,
              style="jenks", title="% Population Aged 16 to 29", 
              palette=mypal9, 
              border.col="gray")+
  tm_borders(col="gray", lwd=0.3)+
  tm_layout(main.title="London Population Aged 16 to 29",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))

#save as pdf
tmap_save(youth, "youth.pdf")

```

## Young People hotspot analysis

```{r}
#define neighbours
msoa_neighbours_fd <- dnearneigh(st_geometry(st_centroid(youth_msoa_sdf)), 0, 3000)


# creates a neighbours list
msoa_spatial_weights_fd <- msoa_neighbours_fd %>%
    nb2listw(., style = "B")


# run the local Gi* test on our  data, note we again use the fd
# weights
youth_LGO <- youth_msoa_sdf %>%
    pull(per_16to29) %>%
    as.vector() %>%
    localG(., msoa_spatial_weights_fd)


# join the local Gi* statistic to `obesity_ward_sdf` spatial dataframe
youth_msoa_sdf <- youth_msoa_sdf %>%
    mutate(youth_LGO_G = as.numeric(youth_LGO))

# create a colour palette
GIColours <- rev(brewer.pal(8, "RdBu"))

#subset london boroughs to include outline
ldnb <- read_sf("./ESRI/London_Borough_Excluding_MHW.shp")
towerhamlets_outline <- subset(ldnb, NAME=="Tower Hamlets")
westminster_outline <- subset(ldnb, NAME=="Westminster")
hackney_outline <- subset(ldnb, NAME=="Hackney")
camden_outline <- subset(ldnb, NAME=="Camden")

# plot the clusters
youth_hotspot <- tm_shape(youth_msoa_sdf) + 
  tm_polygons("youth_LGO_G", style = "pretty", palette=GIColours,
              midpoint=0, title = "Local Gi* statistic") +
  tm_layout(main.title = 'Hot/Cold Spot Map of Youth Population in London', 
            main.title.fontface = 2, fontfamily = "Helvetica",
            legend.outside = TRUE, 
            legend.outside.position = "right", 
            legend.title.size = 1, 
            legend.title.fontface = 2, 
            frame=F) + 
  tm_compass(type = "arrow", position = c("right", "bottom")) +
  tm_scale_bar(breaks = c(0, 5, 10, 15, 20), position = c("left", "bottom"))+
  tm_shape(towerhamlets_outline)+
  tm_borders(col="yellow")+
  tm_shape(westminster_outline)+
  tm_borders(col="yellow")+
  tm_shape(hackney_outline)+
  tm_borders(col="yellow")

#save as pdf
tmap_save(youth_hotspot, "youth_hotspot.pdf")

```

## Education
```{r}
#joining to the spatial dataframe
education_msoa_sdf <- left_join(london_msoa_shape, qualification_msoa, by = c("MSOA11CD" = "msoa_code"))

#defining a custom palette using cartocolors colors
mypal3 <- c("#ede5cf","#e0c2a2","#d39c83","#c1766f","#a65461","#813753","#541f3f")

#no qualification
noqual <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(education_msoa_sdf)+
  tm_fill(col="percent_noqual", n=5,
              style="jenks", title="% Population with no qualification", 
              palette=mypal3, 
              border.col="gray")+
  tm_borders(col="grey", lwd=0.3)+
  tm_layout(main.title="London Population with No Qualification",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))

#save as pdf
tmap_save(noqual, "noqual.pdf")

#define a new palette
mypal4 <- c("#e4f1e1","#b4d9cc","#89c0b6","#63a6a0","#448c8a","#287274","#0d585f")

#Degree and Above map
tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(education_msoa_sdf)+
  tm_fill(col="percent_degreeandabove", n=5,
              style="jenks", title="% Population with a degree and above ", 
              palette=mypal3, 
              border.col="gray")+
  tm_layout(main.title="London Population with a\nDegree and Above",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_borders(col="gray", lwd=0.3)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))
```

## Deprivation

```{r}
#joining to the spatial dataframe
deprivation_msoa_sdf <- left_join(london_msoa_shape, imd_decile_msoa, by = c("MSOA11CD" = "MSOA11CD"))

#define a new palette
mypal5 <- c("#f3cbd3","#eaa9bd","#dd88ac","#ca699d","#b14d8e","#91357d","#6c2167")

#map using pdf
imd <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(deprivation_msoa_sdf)+
  tm_fill(col="mean_imd_decile",
              style="jenks", title="Mean Index of Multiple Deprivation \n (where 10 is the Most Deprived Decile) ", 
              palette=mypal5, 
              border.col="white")+
  tm_borders(col="grey", lwd=0.3)+
  tm_layout(main.title="Mean Index of Multiple Deprivation in London",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))

#save as pdf
tmap_save(imd, "imd.pdf")


```

## Deaths Map

```{r}
#joining to the spatial dataframe
deaths_msoa_sdf <- left_join(london_msoa_shape, london_deaths, by = c("MSOA11CD" = "MSOA11CD"))

#define a new palette
mypal7 <- c("#fef6b5","#ffdd9a","#ffc285","#ffa679","#fa8a76","#f16d7a","#e15383")


#map using pdf
deaths <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(deaths_msoa_sdf)+
  tm_fill(col="covid_19_deaths_per_thousand",
              style="jenks", title="Deaths per 1000 People", 
              palette="Greys", 
              border.col="white")+
  tm_borders(col="grey", lwd=0.3)+
  tm_layout(main.title="COVID-19 Deaths per 1000 People from \n1 March to 17 April 2020",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))

#save as pdf
tmap_save(deaths, "deaths.pdf")
```

## Area of Interest Map

```{r}
#plot the aoi using tmap
 aoi <- tm_shape(ldnb)+
  tm_polygons("white", border.col="grey", lwd=1)+
  tm_text("NAME", size=0.6)+
  tm_layout(main.title="Area of Interest: London,UK",
            main.title.fontface=2,
            fontfamily="Helvetica", 
            legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.title.size=1, 
            legend.title.fontface=2, 
            frame=F)+
  tm_scale_bar(breaks=c(0,5,10,15,20), position=c("left","bottom"))+
  tm_compass(type="arrow", position=c("right", "top"))

#save as pdf
tmap_save(aoi, "aoi_map.pdf")
```


## Bivariate Mapping
```{r}
#clean lsoa data to only include area codes for London
#load london lsoas
lsoa_shape <- read_sf("./ESRI/LSOA_2011_London_gen_MHW.shp")


#joining hospitals data to london lsoas
hospitals_sdf<- inner_join(lsoa_shape,lsoa_data, by = c("LSOA11CD" = "AreaCodes"))


#mapping to see if it worked
hospitals_map <- tm_shape(lsoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(hospitals_sdf)+
  tm_fill(col="hospitalization", n=5,
              style="jenks", title="Hospitalisations", 
              palette=mypal1, 
              border.col="white")+
  tm_borders(col="grey", lwd=0.3)


#first we need to merge the lsoaimd file with the lookup file 
lsoa_hospitals_lookup<- inner_join(msoa_lsoa_lookup,lsoa_data, by = c("LSOA11CD" = "AreaCodes"))


#now aggregate lsoas to msoas (to join with vaccine data)
hospitalizations_msoa <- lsoa_hospitals_lookup %>% group_by(MSOA11CD, MSOA11NM) %>% summarise(mean_hospitalization= mean(hospitalization))

#join msoa now to make an sdf
hospitals_sdf_msoa<- inner_join(london_msoa_shape,hospitalizations_msoa, by = c("MSOA11CD" = "MSOA11CD"))

#mapping to see if it worked
hospitals_map_msoa <- tm_shape(london_msoa_shape)+
  tm_polygons("gray", border.col="gray")+
tm_shape(hospitals_sdf_msoa)+
  tm_fill(col="mean_hospitalization", n=5,
              style="jenks", title="Mean Hospitalisations", 
              palette=mypal1, 
              border.col="white")+
  tm_borders(col="grey", lwd=0.3)

#join with vaccines data
hospitals_and_vaccines_data<- left_join(hospitalizations_msoa,vaccines_msoa_sdf13may, by = c("MSOA11CD" = "MSOA11CD"))

#join msoa now to make an sdf
hospitals_and_vaccines_sdf<- left_join(london_msoa_shape,hospitals_and_vaccines_data, by = c("MSOA11CD" = "MSOA11CD"))


#now make a bivariate map
library(biscale)

# create classes
data <- bi_class(hospitals_and_vaccines_sdf, x = mean_hospitalization, y = per_atleastfirstdose, style = "quantile", dim = 3)

#now make a map
library(ggsn)
# create the basemap
map <- ggplot() +
  geom_sf(data = data, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  labs(
    title = "Bivariate Plot"
  ) +
  bi_theme()

#add scale and compass
m1 <- map+ 
scalebar(data, 
         x.min = -46.85, x.max = -46.38,
         y.min = -24, y.max = -23.35, 
         dist = 10, dist_unit = "km",
             transform = FALSE, model = "WGS84", st.bottom = T)+
  blank()+
  north(data)

#save map as pdf
ggsave("m1.pdf")

#creating the legend
legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher Hospitalization Rate ",
                    ylab = "Higher % Pop. Vaccinated ",
                    size = 8)



# combine map with legend
library(cowplot)
finalPlot <- ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.2, .65, 0.2, 0.2)
```


## KDE
```{r}

#read in the london borough outline
ldnb <- readOGR("ESRI" ,"London_Borough_Excluding_MHW")
plot(ldnb)

#coordinates of all sites
allsites_coords <- read.csv("allsites_coords.csv")

#remove NAs
allsites_coords <- allsites_coords[complete.cases(allsites_coords),]

#convert this to points
allsites_coords %>% st_as_sf(coords = c("Latitude", "Longitude"), crs = 4236) %>%
  # convert into BNG
  st_transform(27700)

allsites_coords <- st_as_sf(allsites_coords, coords=c("Longitude", "Latitude"), crs=4236)

allsites_coords %>% st_set_crs(4326) %>% st_transform(crs=27700) %>% st_crs(27700)
plot(allsites_coords)

#plot on a map
  tm_shape(ldnb)+
  tm_polygons("white", border.col="grey", lwd=1)+
  tm_shape(allsites_coords)+
  tm_dots(col="red")

#first, fix the projection - set everything to 27700
allsites_coords <- st_transform(allsites_coords, "+init=epsg:27700", st_crs(27700))
ldnb <- st_crs(ldnb, 27700)
allsites_coords <- st_crs(allsites_coords, "+init=epsg:27700", st_crs(27700))
ldnb <- st_crs(ldnb, "+init=epsg:27700", st_crs(27700))
projection(ldnb) <- CRS("+init=EPSG:27700")

#checking the reprojections
st_crs(allsites_coords) #now 27700
st_crs(ldnb) #27700


#remove the points which lie outside of London
allsites_coords %>%  st_intersection(ldnb)


#KDE
# create a window of observation
window <- as.owin(ldnb)

# extract coordinates from our bike theft data set as a matrix
vaxsites_xy <- allsites_coords %>%
    st_coordinates()

# create a ppp object
vaxsites_ppp <- ppp(x = vaxsites_xy[, 1], y = vaxsites_xy[, 2], window = window)

# plot our ppp object
plot(vaxsites_ppp)


#dealing with duplicates
# check for duplicates using
anyDuplicated(vaxsites_ppp)

# add an offset to our points using the rjitter function
vaxsites_ppp_jitter <- rjitter(vaxsites_ppp, retry = TRUE, nsim = 1, drop = TRUE)

# count the number of duplicated points of the new jitter data set
anyDuplicated(vaxsites_ppp_jitter)

# plot the resulting ppp object
plot(vaxsites_ppp_jitter)

# kernel density estimation of our bike theft ppp object with 1000m bandwith
plot(density.ppp(vaxsites_ppp_jitter, sigma = 1000))


# create a raster directly from the output of our KDE
kde_1500g_raster <- density.ppp(vaxsites_ppp_jitter, sigma = 600, edge = T) %>%
    raster()

# plot the resulting raster
plot(kde_1000g_raster)


# set the CRS of the kde_400g_raster to BNG
crs(kde_1500g_raster) <- "+init=EPSG:27700"


# map our kde_400g_raster using tmap
kde_map <- tm_shape(kde_1500g_raster) + tm_raster("layer", palette = "YlOrRd")+
   tm_shape(ldnb)+tm_borders(alpha=0.3, col="grey")+
  tm_layout(legend.outside=T, frame=F, main.title="Kernel Density Estimation of \nVaccination Sites in London (As of April 2022)")+
  tm_compass(position=c("right", "top"))+
  tm_scale_bar(breaks=c(5,10,15,20), position=c("left", "bottom"))

#save as pdf
tmap_save(kde_map, "kde_map.pdf")

```


```{r}
```



